This thesis marks the culimation of my journey at MIT---arguably not the
smoothest (or most direct) of sails. While often challenging, my time at MIT
has also been stimulating and broadening, and has caused me to grow in ways
that I am not sure I would have needed to or been able to elsewhere. I am
grateful for the opportunity I have had to deeply study and grow even more
excited about the Internet---a space where the interplay between technology,
policy, and business is incredibly rich and interesting. Participating in
network operator communities, talking with experts directly, reading papers,
writing code, and analyzing data have all contributed to this interest and
excitement, and to this thesis. Relatedly, I am also grateful for this
opportunity to get the CS education that I ``missed'' during my undergraduate
computer engineering career, while also learning about other ways to think
about and address the world, its insitutions, and its challenges through TPP.

While probably contributing to what I perceived to be somewhat rough seas of my
graduate school career here, I should ultimately thank my adivsors Karen and
Dave for giving me the flexibility and freedom to spend my time fully on what
would eventually and somewhat unexpectedly become my thesis research. You might
say I was given enough rope to hang myself, but thankfully, with their help, I
managed to avoid such a fate. With this long leash and the problems that
inevitably occur when doing things for the first time, I gained an appreciation
for what it really means to do research for a living, including the perhaps
most important first steps of picking a good question to answer and being
excited about the work. I also learned how important it is to talk to
people about your work, for feedback, criticism, new ideas, or in the case
of Internet operators, an understanding of how the Internet actually works,
rather than how we might learn about it at the university.

% With regard to this thesis, I am ultimately happy that I settled on a
% data-driven project---it never ceases to be helpful to understand and talk
% about the real world with real data. However, I am shocked at how the clear
% writing of the thesis fails to convey the process undertaken (the hours of
% debugging, hundreds of cpu-hours of processing, or the hundreds of gigabytes
% of data) to ultimately arrive at this point.
%
% Done is better than perfect, and it's hard to be perfect

% ** Include a note about the perversity of establishing causality and reducing
%    bias, as well as doubting your own conclusions **

%More concretely, in terms of this project, I am glad that I got to use data,
%but lament how the presence of so much data (hundreds of gigabytes) diverted
%my attention to building tools to process, manipulate, and anlyze the data
%instead of spending enough time up front writing and presenting the
%information in this thesis. Done is better than perfect, and it's hard to be
%perfect

\vspace{1em}

Finally (last but not least), the obligatory round of thanks for and
acknowledgement of those who influenced me during this effort:

My first and biggest thanks are to Karen Sollins, my adivsor. She first helped
me by hiring me when I was searching for an RA position related to my interests
in the Internet in November of 2008, and has been a constant source of
encouragement, understanding, and good advice on my career as a graduate
student since then. Karen also took efforts to involve me in a number of
interesting projects and has offered much feedback and helpful perspectives on
this work, for which I am greatly appreciative.

I would also like to thank David Clark for his feedback and advice as he became
involved in my project as I changed to focus on what would ultimately become
this thesis. His experience with and insights about Internet policy and
economics together with his technical knoweldge was very helpful, as was his
suggestion to use Elinor Ostrom's CPR framework to think about the Internet
routing table.

Steven Bauer and I first met when I was searching for an RA, but we ultimately
became better friends when he returned to MIT to work on another project. I
would argue that Steve has helped keep me sane, as he has been a great sounding
board, a source of strong feedback, and an advocate, as well as someone to
``geek out'' with on topics such as Python, TCP, and startups. Thanks for all
these things, Steve.

Arthur Berger has been invaluable in helping me plan my investigation and
analysis and offering suggestions and assistance in interpreting the resulting
data, for which I am immensely grateful. Arthur and I also had several
discussions about my professional aspirations that were helpful in my deciding
what to do next, and I appreciate his thoughts and advice to this end.

I would like to thank a number of members of the Internet community for
volunteering to talk with me about some of the various questions I've had about
the Internet. In particular, these individuals' insights and opinions were
extremely valuable in understanding how the Internet works \emph{in practice},
which is often quite different from theory and academic viewpoints. Thank you
Martin Hannigan, Geoff Huston, Tony Li, Bruce Davie, and Jeff Schiller.

Thanks to Susan Perez, the ANA administrative assistant, for always being
incredibly helpful and cheerful. I must acknowledge my officemates in 32-G806
for always providing good company and discussion---especially Jesse Sowell and
Chintan Vaishnav---who have offered research advice and discussion when
asked, and passionate conversation always.

Thanks are due to my parents and sister as well, and in particular my mother,
for their support and encouragement through this process. It started with the
drive down to Massachusetts in 2008 and has continued to the present with
the Calvin and Hobbes comic strip that has graced my inbox daily since April.

Finally, I acknowledge the financial support of Intel, the MIT Communications
Futures Program, and the Natural Sciences and Engineering Research Council of
Canada (NSERC).
